---
title: "Feature Selection"
output: rmarkdown::html_vignette
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(risk3r)
library(broom)
library(dplyr)

data("credit_woe")
```

## Raw Model

Let fit a very raw model to check:

```{r}
train_data <- head(credit_woe, 20000)
test_data  <- tail(credit_woe, 20000)

model_raw <- glm(bad ~ ., family = binomial, data = train_data)

broom::tidy(model_raw)

# some parameter are negative, multicollinearity
broom::tidy(model_raw) %>% filter(estimate <= 0 | p.value > 0.05)

# Yikes!
gg_model_coef(model_raw) 
```

## Satellite Model

```{r, eval=FALSE}
# rf simple
model_rf <- randomForest(bad_good ~ ., data = train_data, do.trace = TRUE, ntree = 500)

metrics(
  train_data$bad_good, 
  predict(model_rf, type = "prob")[,2]
  )
```

## Feature selection via glmnet

From https://glmnet.stanford.edu/articles/glmnet.html

There 2 option for `S`: `lambda.min` and `lambda.1se` , this last option
you have a more regularized model.

This wrapper around the glmnet package take a model as input, then return the
model with the variables non zero from the `glmnet::cv.glmnet()` function 
according with the selected `S` option. This function reorder the variables 
in the same order the coefficient in the glmnet model turn to non zero 
(check the plots when run this funtion).

```{r, message=FALSE}
model_fsglmnet <- featsel_glmnet(model_raw, S = "lambda.1se", trace = TRUE)

broom::tidy(model_fsglmnet)

broom::tidy(model_fsglmnet) %>% filter(estimate <= 0 | p.value > 0.05)
```

## Feature selection via Setpwise forward

This is a wrapper for `stats::step` but the start point model is the null one
`response ~ 1`.

```{r}
model_fsstep <- featsel_stepforward(model_raw, trace = TRUE)

broom::tidy(model_fsstep)

broom::tidy(model_fsstep) %>%
  filter(estimate <= 0 | p.value > 0.05)
```

## Feature selection via repetitive round of drop out loss

From https://ema.drwhy.ai/featureImportance.html

The `stat = "min"` option give the most simpler model because It'more strict:
If any iteration the loss increase then the variable is removed. You can choose
between `c("q25", "mean", "median")`

```{r, eval=FALSE}
model_lss_prmt <- featsel_loss_function_permutations(
  model_raw, 
  stat ="min",
  B = 50
  )

attr(model_lss_prmt, "plots") %>% 
  purrr::reduce(`+`) &
  patchwork::plot_layout(ncol = 1) &
  THEME(base_size = 8)

broom::tidy(model_lss_prmt)

broom::tidy(model_lss_prmt) %>% 
  filter(estimate <= 0 | p.value > 0.05)
```

## Manual

Let suppose you have a nice model, but you want to reduce the number of variables.

Maybe you want check:

```{r, message=FALSE}
model_partials(model_fsstep)

library(ggplot2)

theme_set(theme_minimal(base_size = 7, base_family = ""))

gg_model_partials(model_fsstep) +
    ggplot2::facet_wrap(
      ggplot2::vars(.data$key), ncol = 2,
      scales = "free_y"
      )
```



