---
title: "Simple monotonic WOE retry binning"
output: rmarkdown::html_vignette
editor_options:
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  fig.width = 8,
  fig.height = 5
)
```

```{r setup}
library(risk3r)
library(scorecard)
library(dplyr)
library(ggplot2)

data("germancredit", package = "scorecard")
```

## Motivation

Scorecard development usually needs bins that are easy to explain and stable
across samples. For numeric variables, a monotonic relationship between the bins
and the observed bad rate is often preferred.

`woebin_monotonic()` is intentionally simple. It is a retry wrapper around
`scorecard::woebin()`, not an optimal binning algorithm.

It does not optimize IV under a monotonicity constraint. It does not search all
possible cuts. It only retries numeric variables with coarser binning settings
until the selected metric is monotonic or the retry limits are reached.

The default monotonicity check uses `posprob`. Missing values are kept as a
separate bin and are not included in the monotonicity check. Categorical
variables are passed to `scorecard::woebin()` without forcing monotonicity.

## Methodology

For each numeric variable, the helper follows a basic coarsening strategy:

1. Run `scorecard::woebin()`.
2. Remove the `missing` bin from the monotonicity check.
3. Check whether `posprob` is monotonic across bins.
4. If it is not accepted, increase `count_distr_limit`.
5. If `count_distr_limit` reaches the maximum value, reduce `bin_num_limit`.
6. Keep the first accepted binning.
7. If no accepted binning is found, keep the last valid binning and mark it for
   review in the diagnostic summary.

The result includes a diagnostic table in `attr(bins, "monotonic_summary")`.

## German Credit example

We use a small subset of the German Credit data.

```{r data}
vars <- c(
  "duration.in.month",
  "credit.amount",
  "age.in.years",
  "status.of.existing.checking.account",
  "purpose"
)

dat <- germancredit[, c("creditability", vars)]
```

First, we can run the usual `scorecard::woebin()` process.

```{r regular-bins}
bins_regular <- scorecard::woebin(
  dat,
  y = "creditability",
  x = vars,
  no_cores = 1,
  print_step = 0
)

woebin_summary(bins_regular) %>%
  select(variable, n_categories, iv, monotone, factor)
```

Now we run the simple retry wrapper. By default, `direction = "auto"` accepts
both increasing and decreasing monotonicity. If the expected risk direction is
known, use `direction = "increasing"` or `direction = "decreasing"`.

```{r monotonic-bins}
bins_monotonic <- woebin_monotonic(
  dat,
  y = "creditability",
  x = vars,
  direction = "auto",
  no_cores = 1,
  print_step = 0,
  bin_num_limit = 6,
  verbose = FALSE
)

woebin_summary(bins_monotonic) %>%
  select(variable, n_categories, iv, monotone, factor)
```

The diagnostic summary shows what the helper did for each variable.

```{r monotonic-summary}
attr(bins_monotonic, "monotonic_summary")
```

## Inspect the numeric bins

The helper is meant to make a first monotonic binning attempt. The result still
needs analyst review.

```{r numeric-bin-path}
numeric_vars <- c("duration.in.month", "credit.amount", "age.in.years")

plot_data <- dplyr::bind_rows(bins_monotonic[numeric_vars]) %>%
  filter(breaks != "missing") %>%
  group_by(variable) %>%
  mutate(bin_id = row_number()) %>%
  ungroup()

ggplot(plot_data, aes(bin_id, posprob, group = 1)) +
  geom_line() +
  geom_point() +
  facet_wrap(vars(variable), scales = "free_x") +
  labs(
    x = "Bin order",
    y = "Observed bad rate / posprob",
    title = "Monotonicity check for numeric variables"
  ) +
  theme_minimal()
```

## Important caveat

A monotonic binning result is not automatically a good risk variable.

The analyst still has to validate the direction of risk. For example, a longer
loan duration may be expected to increase risk in many credit products, but that
assumption should be checked against the product, portfolio, policy, and sample
used in the model. Other variables can be monotonic in the data but still have a
risk direction that is hard to justify or unstable out of sample.

In practice, use the monotonic helper as a first pass. Then review:

- whether the direction of `posprob` makes business sense;
- whether the accepted direction matches the expected risk direction;
- whether the number of bins is reasonable;
- whether missing and special-value bins should be kept separate;
- whether the pattern is stable in validation samples or later vintages.

The goal is not to automate judgment away. The goal is to make the first binning
iteration faster and easier to inspect.
