---
title: "Simple monotonic WOE retry binning"
output: rmarkdown::html_vignette
editor_options:
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  fig.width = 8,
  fig.height = 5
)
```

```{r setup}
library(risk3r)
library(scorecard)
library(dplyr)
library(ggplot2)

data("germancredit", package = "scorecard")
```

## Motivation

Scorecard development usually needs bins that are easy to explain and stable
across samples. For numeric variables, a monotonic relationship between the bins
and the observed bad rate is often preferred.

`woebin_monotonic_retry()` is intentionally simple. It is a retry wrapper around
`scorecard::woebin()`, not an optimal binning algorithm.

It does not optimize IV under a monotonicity constraint. It does not search all
possible cuts. It only retries numeric variables with coarser binning settings
until the selected metric is monotonic or the retry limits are reached.

The default monotonicity check uses `posprob`. Missing values are kept as a
separate bin and are not included in the monotonicity check. Categorical
variables are passed to `scorecard::woebin()` without forcing monotonicity.

## Methodology

For each numeric variable, the helper follows a basic coarsening strategy:

1. Run `scorecard::woebin()`.
2. Remove the `missing` bin from the monotonicity check.
3. Check whether `posprob` is monotonic across bins.
4. If it is not monotonic, increase `count_distr_limit`.
5. If `count_distr_limit` reaches the maximum value, reduce `bin_num_limit`.
6. Keep the first monotonic binning.
7. If no monotonic binning is found, keep the last valid binning and mark it for
   review in the diagnostic summary.

The result includes a diagnostic table in `attr(bins, "monotonic_summary")`.
The diagnostic table reports the observed trend for each numeric variable, but
it does not decide whether that trend makes business sense.

## German Credit example

We use a small subset of the German Credit data.

```{r data}
vars <- c(
  "duration.in.month",
  "credit.amount",
  "age.in.years",
  "status.of.existing.checking.account",
  "purpose"
)

dat <- germancredit[, c("creditability", vars)]
```

First, we can run the usual `scorecard::woebin()` process.

```{r regular-bins}
bins_regular <- scorecard::woebin(
  dat,
  y = "creditability",
  x = vars,
  no_cores = 1,
  print_step = 0
)

woebin_summary(bins_regular) %>%
  select(variable, n_categories, iv, monotone, factor)
```

Now we run the simple retry wrapper. It accepts either increasing or decreasing
monotonicity for each numeric variable. This is intentional: the expected risk
direction is variable-specific and should be reviewed after binning.

```{r monotonic-bins}
bins_monotonic <- woebin_monotonic_retry(
  dat,
  y = "creditability",
  x = vars,
  no_cores = 1,
  print_step = 0,
  bin_num_limit = 6,
  verbose = FALSE
)

woebin_summary(bins_monotonic) %>%
  select(variable, n_categories, iv, monotone, factor)
```

The diagnostic summary shows what the helper did for each variable.

```{r monotonic-summary}
attr(bins_monotonic, "monotonic_summary")
```

## Inspect the numeric bins

The helper is meant to make a first monotonic binning attempt. The result still
needs analyst review.

```{r numeric-bin-path}
numeric_vars <- c("duration.in.month", "credit.amount", "age.in.years")

plot_data <- dplyr::bind_rows(bins_monotonic[numeric_vars]) %>%
  filter(breaks != "missing") %>%
  group_by(variable) %>%
  mutate(bin_id = row_number()) %>%
  ungroup()

ggplot(plot_data, aes(bin_id, posprob, group = 1)) +
  geom_line() +
  geom_point() +
  facet_wrap(vars(variable), scales = "free_x") +
  labs(
    x = "Bin order",
    y = "Observed bad rate / posprob",
    title = "Monotonicity check for numeric variables"
  ) +
  theme_minimal()
```

## Important caveat

A monotonic binning result is not automatically a good risk variable.

The analyst still has to validate the direction of risk. For example, a longer
loan duration may be expected to increase risk in many credit products, while
other variables, such as age or income, may reasonably move in another direction.
A single global trend argument would hide that variable-specific judgment.

In practice, use the monotonic helper as a first pass. Then review:

- whether the observed trend of `posprob` makes business sense for each variable;
- whether the number of bins is reasonable;
- whether missing and special-value bins should be kept separate;
- whether the pattern is stable in validation samples or later vintages.

The goal is not to automate judgment away. The goal is to make the first binning
iteration faster and easier to inspect.
