---
title: "Monotonic WOE binning"
output: rmarkdown::html_vignette
editor_options:
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  fig.width = 8,
  fig.height = 5
)
```

```{r setup}
library(risk3r)
library(scorecard)
library(dplyr)
library(ggplot2)

data("germancredit", package = "scorecard")
```

## Motivation

`woebin_monotonic()` separates the orchestration from the numeric binning
method:

- categorical variables always use `scorecard::woebin()`;
- numeric variables use the engine selected with `type`;
- every engine returns one table compatible with an element of a
  `scorecard::woebin()` result.

This makes the method extensible without changing the outer workflow.

Two numeric engines are currently available:

- `type = "scorecard_retry"` uses the original retry/coarsening helper;
- `type = "isoreg"` uses weighted isotonic regression and direct adjacent-bin
  merging.

Neither method replaces analyst judgment. Monotonicity removes irregular
patterns, but the final direction still needs to make sense for the risk
problem.

## Isotonic engine

For one numeric variable, the isotonic engine:

1. Calculates the observed positive rate for every distinct value of `x`.
2. Fits weighted increasing and decreasing isotonic regressions.
3. Applies `count_distr_limit` and `bin_num_limit` to both candidates by merging
   adjacent blocks with the smallest increase in weighted squared error.
4. Selects the final direction with the lowest weighted squared error.
5. Converts the blocks to cut points.
6. Calls `scorecard::woebin()` with those cut points to obtain the usual WOE and
   IV output.

The direction is not imposed in advance. The engine follows the direction that
best fits the observed data after applying the bin constraints.

Missing and special-value bins are excluded from the monotonicity check. Manual
breaks take precedence over either numeric engine.

## German Credit example

```{r data}
vars <- c(
  "duration.in.month",
  "credit.amount",
  "age.in.years",
  "status.of.existing.checking.account",
  "purpose"
)

dat <- germancredit[, c("creditability", vars)]
```

### Retry engine

```{r retry-bins}
bins_retry <- woebin_monotonic(
  dat,
  y = "creditability",
  x = vars,
  type = "scorecard_retry",
  no_cores = 1,
  print_step = 0,
  bin_num_limit = 6,
  count_distr_limit = 0.05,
  count_distr_step = 0.025,
  verbose = FALSE
)

woebin_summary(bins_retry) %>%
  select(variable, n_categories, iv, monotone, factor)
```

### Isotonic engine

```{r isoreg-bins}
bins_isoreg <- woebin_monotonic(
  dat,
  y = "creditability",
  x = vars,
  type = "isoreg",
  no_cores = 1,
  print_step = 0,
  bin_num_limit = 6,
  count_distr_limit = 0.05,
  verbose = FALSE
)

woebin_summary(bins_isoreg) %>%
  select(variable, n_categories, iv, monotone, factor)
```

The isotonic tables keep a few diagnostic attributes:

```{r diagnostics}
attr(bins_isoreg$credit.amount, "monotonic_direction")
attr(bins_isoreg$credit.amount, "isoreg_weighted_sse")
attr(bins_isoreg$credit.amount, "isoreg_breaks")
```

## Inspect the numeric bins

```{r numeric-bin-path}
numeric_vars <- c("duration.in.month", "credit.amount", "age.in.years")

plot_data <- bind_rows(bins_isoreg[numeric_vars]) %>%
  filter(breaks != "missing", !is_special_values) %>%
  group_by(variable) %>%
  mutate(bin_id = row_number()) %>%
  ungroup()

ggplot(plot_data, aes(bin_id, posprob, group = 1)) +
  geom_line() +
  geom_point() +
  facet_wrap(vars(variable), scales = "free_x") +
  labs(
    x = "Bin order",
    y = "Observed bad rate / posprob",
    title = "Monotonic bad rates for numeric variables"
  ) +
  theme_minimal()
```

## Practical review

After the automatic pass, review:

- whether the selected direction makes business sense;
- whether the number and size of bins are useful operationally;
- whether missing and special values should remain separate;
- whether the pattern is stable in validation samples and later vintages;
- whether a manual break is preferable for interpretation or policy reasons.

The goal is to simplify the first binning iteration, not to automate away the
risk decision.
