---
title: "Simple monotonic WOE retry binning"
output: rmarkdown::html_vignette
editor_options:
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  fig.width = 8,
  fig.height = 5
)
```

```{r setup}
library(risk3r)
library(scorecard)
library(dplyr)
library(ggplot2)

data("germancredit", package = "scorecard")
```

## Motivation

`woebin_monotonic_scd_retry()` is intentionally simple. It is a retry wrapper
around `scorecard::woebin()`, not an optimal binning algorithm.

It does not optimize IV under a monotonicity constraint. It does not search all
possible cuts. It only retries numeric variables with coarser binning settings
until `posprob` is monotonic or the retry path is exhausted.

Categorical variables are passed to `scorecard::woebin()` without forcing
monotonicity. Missing values are kept as a separate bin and are not included in
the monotonicity check.

## Methodology

For each numeric variable, the helper follows a basic coarsening strategy:

1. Run `scorecard::woebin()`.
2. Remove the `missing` bin from the monotonicity check.
3. Check whether `posprob` is monotonic across numeric bins.
4. If it is not monotonic, increase `count_distr_limit` by `count_distr_step`.
5. If `count_distr_limit` reaches `0.2`, reduce `bin_num_limit`.
6. Keep the first monotonic binning.
7. If no monotonic binning is found, keep the last valid binning for review.

This is useful as a first pass, but the analyst still has to review whether the
observed direction of risk makes sense for each variable.

## German Credit example

We use a small subset of the German Credit data.

```{r data}
vars <- c(
  "duration.in.month",
  "credit.amount",
  "age.in.years",
  "status.of.existing.checking.account",
  "purpose"
)

dat <- germancredit[, c("creditability", vars)]
```

First, we can run the usual `scorecard::woebin()` process.

```{r regular-bins}
bins_regular <- scorecard::woebin(
  dat,
  y = "creditability",
  x = vars,
  no_cores = 1,
  print_step = 0
)

woebin_summary(bins_regular) %>%
  select(variable, n_categories, iv, monotone, factor)
```

Now we run the simple retry wrapper.

```{r monotonic-bins}
bins_monotonic <- woebin_monotonic_scd_retry(
  dat,
  y = "creditability",
  x = vars,
  no_cores = 1,
  print_step = 0,
  bin_num_limit = 6,
  count_distr_step = 0.025,
  verbose = FALSE
)

woebin_summary(bins_monotonic) %>%
  select(variable, n_categories, iv, monotone, factor)
```

## Inspect the numeric bins

The helper is meant to make a first monotonic binning attempt. The result still
needs analyst review.

```{r numeric-bin-path}
numeric_vars <- c("duration.in.month", "credit.amount", "age.in.years")

plot_data <- dplyr::bind_rows(bins_monotonic[numeric_vars]) %>%
  filter(breaks != "missing") %>%
  group_by(variable) %>%
  mutate(bin_id = row_number()) %>%
  ungroup()

ggplot(plot_data, aes(bin_id, posprob, group = 1)) +
  geom_line() +
  geom_point() +
  facet_wrap(vars(variable), scales = "free_x") +
  labs(
    x = "Bin order",
    y = "Observed bad rate / posprob",
    title = "Monotonicity check for numeric variables"
  ) +
  theme_minimal()
```

## Important caveat

A monotonic binning result is not automatically a good risk variable.

The analyst still has to validate the direction of risk. For example, a longer
loan duration may be expected to increase risk in many credit products, while
other variables, such as age or income, may reasonably move in another direction.
A single global trend argument would hide that variable-specific judgment.

In practice, use the monotonic helper as a first pass. Then review:

- whether the observed trend of `posprob` makes business sense for each variable;
- whether the number of bins is reasonable;
- whether missing and special-value bins should be kept separate;
- whether the pattern is stable in validation samples or later vintages.

The goal is not to automate judgment away. The goal is to make the first binning
iteration faster and easier to inspect.
