---
title: "Monotonic WOE binning"
output: rmarkdown::html_vignette
editor_options:
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  fig.width = 8,
  fig.height = 5
)
```

```{r setup}
library(risk3r)
library(scorecard)
library(dplyr)
library(ggplot2)

data("germancredit", package = "scorecard")
```

## Motivation

`woebin_monotonic()` separates orchestration from the numeric binning method:

- categorical variables always use `scorecard::woebin()`;
- numeric variables use the engine selected with `type`;
- every numeric engine receives one variable and returns one table compatible
  with an element of a `scorecard::woebin()` result.

This makes it possible to add other monotonic methods without changing the
outer workflow.

Two numeric engines are currently available:

- `type = "scorecard_retry"` uses the original retry/coarsening strategy;
- `type = "isoreg"` applies weighted isotonic merging to an initial scorecard
  binning.

Neither method replaces analyst judgment. Monotonicity removes irregular
patterns, but the final direction still needs to make sense for the risk
problem.

## Isotonic engine

For one numeric variable, the isotonic engine:

1. Runs `scorecard::woebin()` with the requested `count_distr_limit`,
   `bin_num_limit`, missing handling, and special values.
2. Orders the regular numeric bins and uses their observed positive rates and
   counts.
3. Fits weighted increasing and decreasing isotonic regressions.
4. Selects the direction with the lowest weighted squared error.
5. Merges adjacent initial bins that belong to the same selected isotonic block.
6. Calls `scorecard::woebin()` again with the resulting cut points to obtain the
   usual WOE and IV table.

Because the isotonic step only merges bins from the initial scorecard result,
it does not create bins smaller than the initial minimum or exceed the initial
maximum number of bins.

The direction is not imposed in advance. The engine follows the direction that
best fits the observed initial-bin rates, while the analyst still decides
whether that direction is economically sensible.

Missing and special-value bins are excluded from the monotonicity fit and
check.

## German Credit example

```{r data}
vars <- c(
  "duration.in.month",
  "credit.amount",
  "age.in.years",
  "status.of.existing.checking.account",
  "purpose"
)

dat <- germancredit[, c("creditability", vars)]
```

### Retry engine

```{r retry-bins}
bins_retry <- woebin_monotonic(
  dat,
  y = "creditability",
  x = vars,
  type = "scorecard_retry",
  no_cores = 1,
  print_step = 0,
  bin_num_limit = 6,
  count_distr_limit = 0.05,
  count_distr_step = 0.025,
  verbose = FALSE
)

woebin_summary(bins_retry) %>%
  select(variable, n_categories, iv, monotone, factor)
```

### Isotonic engine

```{r isoreg-bins}
bins_isoreg <- woebin_monotonic(
  dat,
  y = "creditability",
  x = vars,
  type = "isoreg",
  no_cores = 1,
  print_step = 0,
  bin_num_limit = 6,
  count_distr_limit = 0.05,
  verbose = FALSE
)

woebin_summary(bins_isoreg) %>%
  select(variable, n_categories, iv, monotone, factor)
```

The individual numeric engines can also be used directly when only one
variable is needed:

```{r direct-engine}
credit_amount_bin <- woebin_monotonic_isoreg(
  dat,
  y = "creditability",
  x = "credit.amount",
  no_cores = 1,
  bin_num_limit = 6,
  count_distr_limit = 0.05,
  verbose = FALSE
)

credit_amount_bin
```

## Inspect the numeric bins

```{r numeric-bin-path}
numeric_vars <- c("duration.in.month", "credit.amount", "age.in.years")

plot_data <- bind_rows(bins_isoreg[numeric_vars]) %>%
  filter(breaks != "missing", !is_special_values) %>%
  group_by(variable) %>%
  mutate(bin_id = row_number()) %>%
  ungroup()

ggplot(plot_data, aes(bin_id, posprob, group = 1)) +
  geom_line() +
  geom_point() +
  facet_wrap(vars(variable), scales = "free_x") +
  labs(
    x = "Bin order",
    y = "Observed bad rate / posprob",
    title = "Monotonic bad rates for numeric variables"
  ) +
  theme_minimal()
```

## Practical review

After the automatic pass, review:

- whether the selected direction makes business sense;
- whether the number and size of bins are useful operationally;
- whether missing and special values should remain separate;
- whether the pattern is stable in validation samples and later vintages;
- whether a manual break is preferable for interpretation or policy reasons.

The goal is to simplify the first binning iteration, not to automate away the
risk decision.
